Algebraic File Synchronization: Adequacy and Completeness

### 2017-10-16 p2/works-inverse.tex
Output labels for easier reviewing

\begin{mydef}[$\worksmeqsign$]\label{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (do not break)
all filesystems on which sequences in $\sqs{A}$ are defined.
\end{mydef}

@@

\begin{myclm}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksm{A\cc S}{A}$, that is, if a sequence is defined,
its initial segments are also defined.
\end{myclm}

\begin{myclm}\label{works_restricted}
$\forall \sqs{A},\sqs{B},S: \worksm{\sqs{A}}{\sqs{B}} \Leftarrow \worksm{S\cc \sqs{A}}{S\cc \sqs{B}}$, that is,
the relationship between the domains of the sets $\sqs{A}$ and $\sqs{B}$ continues to hold
even if they are restricted to the image of $S$.
\end{myclm}

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
The proof is by contradiction;
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in two cases.
First, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[(\cxynv\aFS)(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Second, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of values at the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{combine_independent_sequences}
The combination of independent sequences is defined on all filesystems
where both of the sequences are defined:
\[ S\indep T \Rightarrow \worksmbx{S,T}{S\cc T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

@@

\begin{mylem}\label{worksinputmatch}
If $S$ and $T$ are minimal sequences, $\worksmnilb{S,T}$,
and there are commands on node $n$ in both $S$ ($\cxynv\in A$) and $T$ ($\czwnv\in B$),
then the input types of these commands must match, i.e. $[x]=[z]$.
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using a proof by contradiction: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $S$ and $T$ cannot work on the same filesystem.
\end{proof}

@@

\begin{mydef}[canonical mapping, $\T{}$]\label{def_typemapping}
We use $\T{}$ and write $\T{a}$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\T{\vald}=\vald$, $\T{\valf}=\fil$ and $\T{\empt}=\empt$.
\item For filesystems, $\T{\FS}$ is defined by $\T{\FS}(n) = \T{\FS(n)}$.
\item For commands, $\T{\cbrk}=\cbrk$ and $\T{\cxynv} = \cTxynv$.
\item To sequences of commands, the mapping extends as usual: $\T{\emptyseq}=\emptyseq$, and if $S'=\alpha\cc S$, then $\T{S'}=\T{\alpha}\cc \TS$.
\item Finally we extend $\T{}$ to sets of sequences: $\T{\sqs{A}} = \{\TS \whr S\in\sqs{A}\}$.
\end{itemize}
\end{mydef}

@@

As the types of values determine whether a command breaks a filesystem,
we know $\Dom{A}=\Dom{\TA}$ and therefore $\Dom{\sqs{A}}=\Dom{\T{\sqs{A}}}$.
From this it follows that
% TODO what's the set of all filesystems?
% TODO:
\begin{mycor}\label{repr_works_is_same}
\begin{gather*}
\worksmnil{\sqs{A}} \Longleftrightarrow \worksmnil{\T{\sqs{A}}} \\
\worksm{\sqs{A}}{\sqs{B}} \Longleftrightarrow \worksm{\T{\sqs{A}}}{\T{\sqs{B}}}.
\end{gather*}
\end{mycor}

@@

\begin{mylem}\label{repr_comm_inject}
Canonical commands $\T{\alpha}$ are partial bijections over canonical filesystems
($\T{\FS}$, $\T{\GS}$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that canonical functions are injective except for cases
where they are not defined:
\begin{align*}
\forall\FS,\GS,\alpha:& \\
& \T{\FS}\neq\T{\GS} \wedge \T{\alpha}(\T{\FS})\neq\fsbroken \\
& \Rightarrow \T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS}).
\end{align*}
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\T{\alpha} = \cTxynv$.
If $\T{\FS}$ and $\T{\GS}$ have different values (types) at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\T{\FS}$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\T{\FS})\neq\fsbroken$, we know $[\FS(n)]=[x]$.
We also know that $\T{\GS}(n)\neq\T{\FS}(n)$, and therefore
$[\GS(n)]\neq[x]$.
This implies $\cTxynv(\T{\GS})=\fsbroken$,
and again we have $\T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS})$.
\end{proof}

@@

\begin{mylem}\label{r_invmove}
A common initial segment of a set of canonical sequences can be moved 
to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\ia}{B\cc\ic}
\ \Longleftrightarrow \ \worksm{\ibi\cc\ia}{\ic} 
\ \wedge \  \worksm{\ia}{B}
\end{gather*}
\end{mylem}
\begin{proof}
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

@@

\begin{mylem}\label{indep_prefix_combine}
The combination of two sequences with a common head and independent tails 
is defined wherever the two sequences are defined:
% Use A\cc\{B,C\}
\[ B\indep C \Rightarrow \worksmbx{A\cc B, A\cc C}{A\cc B\cc C} \]
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnil{A}$, so we assume that that is the case.
From $B\indep C$ and \cref{combine_independent_sequences} we know
$\worksmbx{B, C}{B\cc C}$.
Based on \cref{repr_works_is_same} this is equivalent to
$\worksmbx{\TB, \TC}{\TB\cc \TC}$.
Since $\TA^{-1}\cc \TA \equiv \emptyseq$, this can also be written as
\[\worksm{\TA^{-1}\cc\{\TA\cc \TB, \TA\cc \TC\}}{\TB\cc \TC}.\]
From \cref{worksextpostfix} we also know that
$\worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA}$.
Combining the last two statements using \cref{r_invmove}
we get
\[ \worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA\cc \TB\cc \TC}, \]
which with \cref{repr_works_is_same} proves the
\namecref{indep_prefix_combine}.
\end{proof}

>>>>

\begin{mydef}[$\worksmeqsign$]\mlabel{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (do not break)
all filesystems on which sequences in $\sqs{A}$ are defined.
\end{mydef}

@@

\begin{myclm}\mlabel{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksm{A\cc S}{A}$, that is, if a sequence is defined,
its initial segments are also defined.
\end{myclm}

\begin{myclm}\mlabel{works_restricted}
$\forall \sqs{A},\sqs{B},S: \worksm{\sqs{A}}{\sqs{B}} \Rightarrow \worksm{S\cc \sqs{A}}{S\cc \sqs{B}}$, that is,
the relationship between the domains of the sets $\sqs{A}$ and $\sqs{B}$ continues to hold
even if they are restricted to the image of $S$.
\end{myclm}

@@

\begin{myax}\mlabel{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
The proof is by contradiction;
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in two cases.
First, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[(\cxynv\aFS)(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Second, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of values at the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\mlabel{combine_independent_sequences}
The combination of independent sequences is defined on all filesystems
where both of the sequences are defined:
\[ S\indep T \Rightarrow \worksmbx{S,T}{S\cc T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

@@

\begin{mylem}\mlabel{worksinputmatch}
If $S$ and $T$ are minimal sequences, $\worksmnilb{S,T}$,
and there are commands on node $n$ in both $S$ ($\cxynv\in A$) and $T$ ($\czwnv\in B$),
then the input types of these commands must match, i.e. $[x]=[z]$.
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using a proof by contradiction: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $S$ and $T$ cannot work on the same filesystem.
\end{proof}

@@

\begin{mydef}[canonical mapping, $\T{}$]\mlabel{def_typemapping}
We use $\T{}$ and write $\T{a}$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\T{\vald}=\vald$, $\T{\valf}=\fil$ and $\T{\empt}=\empt$.
\item For filesystems, $\T{\FS}$ is defined by $\T{\FS}(n) = \T{\FS(n)}$.
\item For commands, $\T{\cbrk}=\cbrk$ and $\T{\cxynv} = \cTxynv$.
\item To sequences of commands, the mapping extends as usual: $\T{\emptyseq}=\emptyseq$, and if $S'=\alpha\cc S$, then $\T{S'}=\T{\alpha}\cc \TS$.
\item Finally we extend $\T{}$ to sets of sequences: $\T{\sqs{A}} = \{\TS \whr S\in\sqs{A}\}$.
\end{itemize}
\end{mydef}

@@

As the types of values determine whether a command breaks a filesystem,
we know $\Dom{A}=\Dom{\TA}$ and therefore $\Dom{\sqs{A}}=\Dom{\T{\sqs{A}}}$.
From this it follows that
% TODO what's the set of all filesystems?
% TODO:
\begin{mycor}\mlabel{repr_works_is_same}
\begin{gather*}
\worksmnil{\sqs{A}} \Longleftrightarrow \worksmnil{\T{\sqs{A}}} \\
\worksm{\sqs{A}}{\sqs{B}} \Longleftrightarrow \worksm{\T{\sqs{A}}}{\T{\sqs{B}}}.
\end{gather*}
\end{mycor}

@@

\begin{mylem}\mlabel{repr_comm_inject}
Canonical commands $\T{\alpha}$ are partial bijections over canonical filesystems
($\T{\FS}$, $\T{\GS}$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that canonical functions are injective except for cases
where they are not defined:
\begin{align*}
\forall\FS,\GS,\alpha:& \\
& \T{\FS}\neq\T{\GS} \wedge \T{\alpha}(\T{\FS})\neq\fsbroken \\
& \Rightarrow \T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS}).
\end{align*}
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\T{\alpha} = \cTxynv$.
If $\T{\FS}$ and $\T{\GS}$ have different values (types) at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\T{\FS}$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\T{\FS})\neq\fsbroken$, we know $[\FS(n)]=[x]$.
We also know that $\T{\GS}(n)\neq\T{\FS}(n)$, and therefore
$[\GS(n)]\neq[x]$.
This implies $\cTxynv(\T{\GS})=\fsbroken$,
and again we have $\T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS})$.
\end{proof}

@@

\begin{mylem}\mlabel{r_invmove}
A common initial segment of a set of canonical sequences can be moved 
to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\ia}{B\cc\ic}
\ \Longleftrightarrow \ \worksm{\ibi\cc\ia}{\ic} 
\ \wedge \  \worksm{\ia}{B}
\end{gather*}
\end{mylem}
\begin{proof}
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

@@

\begin{mylem}\mlabel{indep_prefix_combine}
The combination of two sequences with a common head and independent tails 
is defined wherever the two sequences are defined:
% Use A\cc\{B,C\}
\[ B\indep C \Rightarrow \worksmbx{A\cc B, A\cc C}{A\cc B\cc C} \]
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnil{A}$, so we assume that that is the case.
From $B\indep C$ and \cref{combine_independent_sequences} we know
$\worksmbx{B, C}{B\cc C}$.
Based on \cref{repr_works_is_same} this is equivalent to
$\worksmbx{\TB, \TC}{\TB\cc \TC}$.
Since $\TA^{-1}\cc \TA \equiv \emptyseq$, this can also be written as
\[\worksm{\TA^{-1}\cc\{\TA\cc \TB, \TA\cc \TC\}}{\TB\cc \TC}.\]
From \cref{worksextpostfix} we also know that
$\worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA}$.
Combining the last two statements using \cref{r_invmove}
we get
\[ \worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA\cc \TB\cc \TC}, \]
which with \cref{repr_works_is_same} proves the
\namecref{indep_prefix_combine}.
\end{proof}
